### 2022-01-17 Non_Identical_r0/Rmarkdown/Classical/OSMAC_Model_Free_Method.Rmd

---
title: "OSMAC_Model_Free_Method; Model: `r params$Model_Path`; "
author: "M.Amalan"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,fig.height = 9,fig.width = 9,message = FALSE,warning=FALSE)
library(ggplot2)
library(here)
library(dplyr)
library(tidyr)
library(gganimate)
library(LaplacesDemon)
library(linguisticsdown)
library(gganimate)
library(viridis)
library(compiler)
library(data.table)
library(ggpubr)

# Cordeiro Bias Estimation ----
Cordeiro<-function(XData,With_bias)
{
  p <- as.vector(exp(XData%*%as.vector(With_bias)))
  W <- diag(p)
  inverse_term <- solve(t(XData)%*%W%*%XData)

  Term1 <- inverse_term%*%t(XData)%*%W
  Term2 <- diag(diag(XData%*%(inverse_term)%*%t(XData))) %*% rep(-0.5,nrow(XData))

  bias <- as.vector(Term1%*%Term2)
  return(bias)
}

Cordeiro<-cmpfun(Cordeiro)

FullDataPath<-here("Non_Identical_r0","Generate_Big_Data",params$Model_Path,"No_Correlated_Covariate.RData")
load(here(FullDataPath))

Full_model<-glm(Y~.-1,data = as.data.frame(Simulated_Data[[2]]$Real_Model),family = "poisson")
Real_Vline<-data.frame("Theta"=c(paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1]))),
                       "Values"=Simulated_Data$Basic$Theta)
Estimated_Vline<-data.frame("Theta"=c(paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1]))),
                            "Values"=Full_model$coefficients)
Bias_Vline<-data.frame("Theta"=c(paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1]))),"Values"=0)
```

# OSMAC ;  Response Model : `r params$Model_Path`

## Real Model Data

```{r Load Data}
# Full Data RM
Full_model<-glm(Y~.-1,data = as.data.frame(Simulated_Data[[2]]$Real_Model),family = "poisson")
Estimated_Vline_RM<-data.frame("Theta"=c(paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1]))),
                               "Values"=Full_model$coefficients)
Full_Est_Parameter_RM<-as.data.frame(t(Estimated_Vline_RM$Values))
Full_Est_Bias_RM<-as.data.frame(t(Cordeiro(XData = Simulated_Data[[2]]$Real_Model[,-1],
                                           With_bias = as.matrix(Full_Est_Parameter_RM))))
colnames(Full_Est_Parameter_RM)<-colnames(Full_Est_Bias_RM)<-c(paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))
Full_Bias_RM<-sqrt(sum(Full_Est_Bias_RM^2))
Temp_P<-exp(Simulated_Data[[2]]$Real_Model[,-1]%*% t(as.matrix(Full_Est_Parameter_RM)))
Temp_W<-diag(as.vector(Temp_P))
Temp_Info<-t(Simulated_Data[[2]]$Real_Model[,-1])%*% Temp_W %*% Simulated_Data[[2]]$Real_Model[,-1]
Full_Variance_RM<-tr(solve(Temp_Info))
Full_Information_RM<-det(Temp_Info)

Path<-here("Non_Identical_r0","Outputs","Classical",params$Model_Path,"OSMAC_Model_Free/")
load(here(Path,"OSMAC_MF_output.RData"))

Est_Parameter_RM<-rbind.data.frame(Results_OSMAC_1$mMSE_Output,
                                   Results_OSMAC_1$mVc_Output)

Est_Bias_RM<-rbind.data.frame(Bias_OSMAC_1$mMSE_Output,
                              Bias_OSMAC_1$mVc_Output)

Utility_RM<-rbind.data.frame(Utility_OSMAC_1$mMSE_Output,
                             Utility_OSMAC_1$mVc_Output)

colnames(Est_Parameter_RM)<-colnames(Est_Bias_RM)<-c("Type","Subsample_Size",
                                                     paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))
colnames(Utility_RM)<-c("Type","Subsample_Size","A_optimality","D_optimality")

remove(Results_OSMAC_1,Bias_OSMAC_1,Utility_OSMAC_1)
```

## Parameters

```{r Animated Plots RM}
Est_Parameter_RM %>%
  pivot_longer(starts_with("Theta"), names_to = "Theta", values_to = "Estimated") %>%
ggplot(., aes(x = Estimated, group = as.factor(Theta), frame=Subsample_Size)) + 
  geom_histogram(alpha=0.8)+
  facet_grid(Type~Theta,scales = "free_x",shrink = TRUE)+
  geom_vline(data= Estimated_Vline,aes(xintercept = Values,color="red"),linetype='dashed',size=1)+
  scale_color_discrete(name = "Parameter values", labels = c("Estimated"))+
  labs(title=paste0("Real Model :",params$Model_Path," Data : Estimated Theta"),
      subtitle = "Sub-sample Size: {current_frame}")+
  transition_manual(Subsample_Size)+
  theme_light()+ylab("Frequency")+xlab("Estimates")->para_1

anim_save(filename="para_1.gif",animation = para_1,
          path=here("Non_Identical_r0","htmloutputs","Classical",params$Model_Path,"OSMAC_Model_Free","Plots"))

include_graphics2(here("Non_Identical_r0","htmloutputs","Classical",params$Model_Path,"OSMAC_Model_Free","Plots","para_1.gif"))
```

## Bias

```{r Animated Plots for Bias RM}
Est_Bias_RM %>%
  pivot_longer(starts_with("Theta"), names_to = "Theta", values_to = "Estimated") %>%
ggplot(., aes(x = Estimated, group = as.factor(Theta), frame=Subsample_Size)) + 
  geom_histogram(alpha=0.8)+
  facet_grid(Type~Theta,scales = "free_x",shrink = TRUE)+
  geom_vline(data= Bias_Vline,aes(xintercept = Values,color="red"),linetype='dashed',size=1)+
  scale_color_discrete(name = "Bias values", labels = c("Real"))+
  labs(title=paste0("Real Model :",params$Model_Path," Data : Estimated Bias"),
      subtitle = "Sub-sample Size: {current_frame}")+
  transition_manual(Subsample_Size)+theme_light()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  ylab("Frequency")+xlab("Estimates")->Bias_1

anim_save(filename="Bias_1.gif",animation = Bias_1,
          path=here("Non_Identical_r0","htmloutputs","Classical",params$Model_Path,"OSMAC_Model_Free","Plots"))

include_graphics2(here("Non_Identical_r0","htmloutputs","Classical",params$Model_Path,"OSMAC_Model_Free","Plots","Bias_1.gif"))
```

## Utility values

```{r Animated Utility plots RM,fig.width=7,fig.height=7}
Utility_RM %>%
  ggplot(.,aes(x=log(A_optimality)))+
  geom_histogram()+facet_wrap(~Type)+
  theme_light()+ylab("Frequency")+xlab("log(A-optimality)")+
  transition_states(Subsample_Size)+
  ggtitle("Real Model : Subsample : {closest_state}")->U_1

anim_save(filename="U_1.gif",animation = U_1,
          path=here("Non_Identical_r0","htmloutputs","Classical",params$Model_Path,"OSMAC_Model_Free","Plots"))

include_graphics2(here("Non_Identical_r0","htmloutputs","Classical",params$Model_Path,"OSMAC_Model_Free","Plots","U_1.gif"))

Utility_RM %>%
  ggplot(.,aes(x=log(D_optimality)))+
  geom_histogram()+facet_wrap(~Type)+
  theme_light()+ylab("Frequency")+xlab("log(D-optimality) ")+
  transition_states(Subsample_Size)+
  ggtitle("Real Model : Subsample : {closest_state}")->U_2

anim_save(filename="U_2.gif",animation = U_2,
          path=here("Non_Identical_r0","htmloutputs","Classical",params$Model_Path,"OSMAC_Model_Free","Plots"))

include_graphics2(here("Non_Identical_r0","htmloutputs","Classical",params$Model_Path,"OSMAC_Model_Free","Plots","U_2.gif"))
```

# Assumed Model 1

```{r Load the Data AM 1}
# Full Data AM 1
Full_model<-glm(Y~.-1,data = as.data.frame(Simulated_Data[[2]]$Assumed_Model_1),family = "poisson")
Estimated_Vline_AM_1<-data.frame("Theta"=c(paste0("Theta",0:(ncol(Simulated_Data[[2]]$Assumed_Model_1)-2))),
                                 "Values"=Full_model$coefficients)
Full_Est_Parameter_AM_1<-as.data.frame(t(Estimated_Vline_AM_1$Values))
Full_Est_Bias_AM_1<-as.data.frame(t(Cordeiro(XData = Simulated_Data[[2]]$Assumed_Model_1[,-1],
                                      With_bias = as.matrix(Full_Est_Parameter_AM_1))))
colnames(Full_Est_Parameter_AM_1)<-colnames(Full_Est_Bias_AM_1)<-c(paste0("Theta",0:(ncol(Simulated_Data[[2]]$Assumed_Model_1)-2)))

Full_Bias_AM_1<-sqrt(sum(Full_Est_Bias_AM_1^2))
Temp_P<-exp(Simulated_Data[[2]]$Assumed_Model_1[,-1]%*% t(as.matrix(Full_Est_Parameter_AM_1)))
Temp_W<-diag(as.vector(Temp_P))
Temp_Info<-t(Simulated_Data[[2]]$Assumed_Model_1[,-1])%*% Temp_W %*% Simulated_Data[[2]]$Assumed_Model_1[,-1]
Full_Variance_AM_1<-tr(solve(Temp_Info))
Full_Information_AM_1<-det(Temp_Info)

Est_Parameter_AM_1<-rbind.data.frame(Results_OSMAC_2$mMSE_Output,
                                     Results_OSMAC_2$mVc_Output)

Est_Bias_AM_1<-rbind.data.frame(Bias_OSMAC_2$mMSE_Output,
                                Bias_OSMAC_2$mVc_Output)

Utility_AM_1<-rbind.data.frame(Utility_OSMAC_2$mMSE_Output,
                               Utility_OSMAC_2$mVc_Output)

colnames(Est_Parameter_AM_1)<-colnames(Est_Bias_AM_1)<-
  c("Type","Subsample_Size",paste0("Theta",0:(ncol(Simulated_Data[[2]]$Assumed_Model_1)-2)))
colnames(Utility_AM_1)<-c("Type","Subsample_Size","A_optimality","D_optimality")

remove(Results_OSMAC_2,Bias_OSMAC_2,Utility_OSMAC_2)

```

# Assumed Model 2

```{r Load the Data AM 2}
# Full Data AM 2
Full_model<-glm(Y~.-1,data = as.data.frame(Simulated_Data[[2]]$Assumed_Model_2),family = "poisson")
Estimated_Vline_AM_2<-data.frame("Theta"=c(paste0("Theta",0:(ncol(Simulated_Data[[2]]$Assumed_Model_2)-2))),
                                 "Values"=Full_model$coefficients)
Full_Est_Parameter_AM_2<-as.data.frame(t(Estimated_Vline_AM_2$Values))
Full_Est_Bias_AM_2<-as.data.frame(t(Cordeiro(XData = Simulated_Data[[2]]$Assumed_Model_2[,-1],
                                      With_bias = as.matrix(Full_Est_Parameter_AM_2))))
colnames(Full_Est_Parameter_AM_2)<-colnames(Full_Est_Bias_AM_2)<-c(paste0("Theta",0:(ncol(Simulated_Data[[2]]$Assumed_Model_2)-2)))

Full_Bias_AM_2<-sqrt(sum(Full_Est_Bias_AM_2^2))
Temp_P<-exp(Simulated_Data[[2]]$Assumed_Model_2[,-1]%*% t(as.matrix(Full_Est_Parameter_AM_2)))
Temp_W<-diag(as.vector(Temp_P))
Temp_Info<-t(Simulated_Data[[2]]$Assumed_Model_2[,-1])%*% Temp_W %*% Simulated_Data[[2]]$Assumed_Model_2[,-1]
Full_Variance_AM_2<-tr(solve(Temp_Info))
Full_Information_AM_2<-det(Temp_Info)

Est_Parameter_AM_2<-rbind.data.frame(Results_OSMAC_3$mMSE_Output,
                                     Results_OSMAC_3$mVc_Output)

Est_Bias_AM_2<-rbind.data.frame(Bias_OSMAC_3$mMSE_Output,
                                Bias_OSMAC_3$mVc_Output)

Utility_AM_2<-rbind.data.frame(Utility_OSMAC_3$mMSE_Output,
                               Utility_OSMAC_3$mVc_Output)

colnames(Est_Parameter_AM_2)<-colnames(Est_Bias_AM_2)<-
  c("Type","Subsample_Size",paste0("Theta",0:(ncol(Simulated_Data[[2]]$Assumed_Model_2)-2)))
colnames(Utility_AM_2)<-c("Type","Subsample_Size","A_optimality","D_optimality")

remove(Results_OSMAC_3,Bias_OSMAC_3,Utility_OSMAC_3)

```

# Assumed Model 3

```{r Load the Data AM 3}
# Full Data AM 3
Full_model<-glm(Y~.-1,data = as.data.frame(Simulated_Data[[2]]$Assumed_Model_3),family = "poisson")
Estimated_Vline_AM_3<-data.frame("Theta"=c(paste0("Theta",0:(ncol(Simulated_Data[[2]]$Assumed_Model_3)-2))),
                                 "Values"=Full_model$coefficients)
Full_Est_Parameter_AM_3<-as.data.frame(t(Estimated_Vline_AM_3$Values))
Full_Est_Bias_AM_3<-as.data.frame(t(Cordeiro(XData = Simulated_Data[[2]]$Assumed_Model_3[,-1],
                                      With_bias = as.matrix(Full_Est_Parameter_AM_3))))
colnames(Full_Est_Parameter_AM_3)<-colnames(Full_Est_Bias_AM_3)<-c(paste0("Theta",0:(ncol(Simulated_Data[[2]]$Assumed_Model_3)-2)))

Full_Bias_AM_3<-sqrt(sum(Full_Est_Bias_AM_3^2))
Temp_P<-exp(Simulated_Data[[2]]$Assumed_Model_3[,-1]%*% t(as.matrix(Full_Est_Parameter_AM_3)))
Temp_W<-diag(as.vector(Temp_P))
Temp_Info<-t(Simulated_Data[[2]]$Assumed_Model_3[,-1])%*% Temp_W %*% Simulated_Data[[2]]$Assumed_Model_3[,-1]
Full_Variance_AM_3<-tr(solve(Temp_Info))
Full_Information_AM_3<-det(Temp_Info)

Est_Parameter_AM_3<-rbind.data.frame(Results_OSMAC_4$mMSE_Output,
                                     Results_OSMAC_4$mVc_Output)

Est_Bias_AM_3<-rbind.data.frame(Bias_OSMAC_4$mMSE_Output,
                                Bias_OSMAC_4$mVc_Output)

Utility_AM_3<-rbind.data.frame(Utility_OSMAC_4$mMSE_Output,
                               Utility_OSMAC_4$mVc_Output)

colnames(Est_Parameter_AM_3)<-colnames(Est_Bias_AM_3)<-
  c("Type","Subsample_Size",paste0("Theta",0:(ncol(Simulated_Data[[2]]$Assumed_Model_3)-2)))
colnames(Utility_AM_3)<-c("Type","Subsample_Size","A_optimality","D_optimality")

remove(Results_OSMAC_4,Bias_OSMAC_4,Utility_OSMAC_4)

```

# Comparison for Efficiency

```{r Comparison of Summary Data}
Final_MSE<-function(Real,Estimated_1,Estimated_2,Estimated_3,Estimated_4,Estimated_5,Model)
{
  For_Theta0<-sum((Real[1]-Estimated_1)^2,na.rm = TRUE)
  For_Theta1<-sum((Real[2]-Estimated_2)^2,na.rm = TRUE)
  For_Theta2<-sum((Real[3]-Estimated_3)^2,na.rm = TRUE)

  if(Model=="Model_1")
  {
    return((For_Theta0+For_Theta1+For_Theta2)/length(Estimated_1))
  }  
  if(Model=="Model_2" || Model=="Model_3")
  {
    For_Theta3<-sum((Real[4]-Estimated_4)^2,na.rm = TRUE)
    return((For_Theta0+For_Theta1+For_Theta2+For_Theta3)/length(Estimated_1))
  }
  if(Model=="Model_4")
  {
    For_Theta3<-sum((Real[4]-Estimated_4)^2,na.rm = TRUE)
    For_Theta4<-sum((Real[5]-Estimated_5)^2,na.rm = TRUE)
    return((For_Theta0+For_Theta1+For_Theta2+For_Theta3+For_Theta4)/length(Estimated_1))
  }
}

# Combine datasets
Est_Utility<-rbind(data.frame("Model"="Real Model",Utility_RM),
                   data.frame("Model"="Assumed Model 1",Utility_AM_1),
                   data.frame("Model"="Assumed Model 2",Utility_AM_2),
                   data.frame("Model"="Assumed Model 3",Utility_AM_3))

remove(Utility_AM_3,Utility_AM_2,Utility_AM_1,Utility_RM,
       Bias_Vline,Temp_Info,Temp_W,Temp_P,
       Estimated_Vline,Full_model)

# Summarise A- optimality
Est_Utility %>%
  dplyr::select(Type,Model,Subsample_Size,A_optimality) %>%
  group_by(Type,Model,Subsample_Size) %>%
  mutate(mean_Variance=mean(A_optimality))%>%
  distinct(mean_Variance) %>% 
  forestmangr::round_df(8)->All_VAR

# Summarise D- optimality
Est_Utility %>%
  dplyr::select(Type,Model,Subsample_Size,D_optimality) %>%
  group_by(Type,Model,Subsample_Size) %>%
  mutate(mean_Info=mean(D_optimality))%>%
  distinct(mean_Info) %>% 
  forestmangr::round_df(8)->All_Info

# Model 1
if(params$Model_Path=="Model_1")
{
  # Summarize SMSE
  Est_Parameter_RM %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2) %>%
    group_by(Type,Subsample_Size) %>%
    mutate( MSE=Final_MSE(Real = Real_Vline$Values,
                            Estimated_1=Theta0, 
                            Estimated_2=Theta1,
                            Estimated_3=Theta2,
                            Model = params$Model_Path ))%>%
    distinct(MSE) %>% 
    forestmangr::round_df(8)->All_MSE_RM

  Est_Parameter_AM_1 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate( MSE=Final_MSE(Real = Estimated_Vline_AM_1$Values,
                            Estimated_1=Theta0, 
                            Estimated_2=Theta1,
                            Estimated_3=Theta2,
                            Estimated_4=Theta3,
                            Model = "Model_2" ))%>%
    distinct(MSE) %>% 
    forestmangr::round_df(8)->All_MSE_AM_1

  Est_Parameter_AM_2 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate( MSE=Final_MSE(Real = Estimated_Vline_AM_2$Values,
                            Estimated_1=Theta0, 
                            Estimated_2=Theta1,
                            Estimated_3=Theta2,
                            Estimated_4=Theta3,
                            Model = "Model_3" ))%>%
    distinct(MSE) %>% 
    forestmangr::round_df(8)->All_MSE_AM_2

  Est_Parameter_AM_3 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3,Theta4) %>%
    group_by(Type,Subsample_Size) %>%
    mutate( MSE=Final_MSE(Real = Estimated_Vline_AM_3$Values,
                            Estimated_1=Theta0, 
                            Estimated_2=Theta1,
                            Estimated_3=Theta2,
                            Estimated_4=Theta3,
                            Estimated_5=Theta4,
                            Model = "Model_4" ))%>%
    distinct(MSE) %>% 
    forestmangr::round_df(8)->All_MSE_AM_3

  All_MSE<-rbind(data.frame("Model"="Real Model",All_MSE_RM),
                 data.frame("Model"="Assumed Model 1",All_MSE_AM_1),
                 data.frame("Model"="Assumed Model 2",All_MSE_AM_2),
                 data.frame("Model"="Assumed Model 3",All_MSE_AM_3))

  remove(All_MSE_RM,All_MSE_AM_1,All_MSE_AM_2,All_MSE_AM_3)

  # Summarise Bias
  Est_Bias_RM %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + 
                                 Theta2*Theta2)) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(8)->All_Bias_RM

  Est_Bias_AM_1 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + 
                                 Theta2*Theta2+ Theta3*Theta3)) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(8)->All_Bias_AM_1

   Est_Bias_AM_2 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + 
                                 Theta2*Theta2+ Theta3*Theta3)) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(8)->All_Bias_AM_2

  Est_Bias_AM_3 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3,Theta4) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + 
                                 Theta2*Theta2+ Theta3*Theta3+ Theta4*Theta4)) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(8)->All_Bias_AM_3

  All_Bias<-rbind(data.frame("Model"="Real Model",All_Bias_RM),
                  data.frame("Model"="Assumed Model 1",All_Bias_AM_1),
                  data.frame("Model"="Assumed Model 2",All_Bias_AM_2),
                  data.frame("Model"="Assumed Model 3",All_Bias_AM_3))

  remove(All_Bias_AM_1,All_Bias_AM_2,All_Bias_AM_3,All_Bias_RM)

  # Summarise Bias = sqrt(Bias^2)
  Est_Bias_RM %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 +
                            Theta2*Theta2) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(4)->All_Bias_squared_RM

  Est_Bias_AM_1 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 + 
                            Theta2*Theta2 + Theta3*Theta3) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(4)->All_Bias_squared_AM_1

  Est_Bias_AM_2 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 + 
                            Theta2*Theta2 + Theta3*Theta3) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(4)->All_Bias_squared_AM_2

  Est_Bias_AM_3 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3,Theta4) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 + 
                            Theta2*Theta2 + Theta3*Theta3 + Theta4*Theta4) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(4)->All_Bias_squared_AM_3

  All_Bias_squared<-rbind(data.frame("Model"="Real Model",All_Bias_squared_RM),
                          data.frame("Model"="Assumed Model 1",All_Bias_squared_AM_1),
                          data.frame("Model"="Assumed Model 2",All_Bias_squared_AM_2),
                          data.frame("Model"="Assumed Model 3",All_Bias_squared_AM_3))

  remove(All_Bias_squared_RM,All_Bias_squared_AM_1,All_Bias_squared_AM_2,All_Bias_squared_AM_3)
}

# Model 2 and Model 3
if(params$Model_Path=="Model_2" || params$Model_Path=="Model_3")
{
  # Summarize SMSE
  Est_Parameter_RM %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate( MSE=Final_MSE(Real = Real_Vline$Values,
                            Estimated_1=Theta0, 
                            Estimated_2=Theta1,
                            Estimated_3=Theta2,
                            Estimated_4=Theta3,
                            Model = params$Model_Path ))%>%
    distinct(MSE) %>% 
    forestmangr::round_df(8)->All_MSE_RM

  Est_Parameter_AM_1 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2) %>%
    group_by(Type,Subsample_Size) %>%
    mutate( MSE=Final_MSE(Real = Estimated_Vline_AM_1$Values,
                            Estimated_1=Theta0, 
                            Estimated_2=Theta1,
                            Estimated_3=Theta2,
                            Model = "Model_1" ))%>%
    distinct(MSE) %>% 
    forestmangr::round_df(8)->All_MSE_AM_1

  Est_Parameter_AM_2 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate( MSE=Final_MSE(Real = Estimated_Vline_AM_2$Values,
                            Estimated_1=Theta0, 
                            Estimated_2=Theta1,
                            Estimated_3=Theta2,
                            Estimated_4=Theta3,
                            Model = "Model_3" ))%>%
    distinct(MSE) %>% 
    forestmangr::round_df(8)->All_MSE_AM_2

  Est_Parameter_AM_3 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3,Theta4) %>%
    group_by(Type,Subsample_Size) %>%
    mutate( MSE=Final_MSE(Real = Estimated_Vline_AM_3$Values,
                            Estimated_1=Theta0, 
                            Estimated_2=Theta1,
                            Estimated_3=Theta2,
                            Estimated_4=Theta3,
                            Estimated_5=Theta4,
                            Model = "Model_4" ))%>%
    distinct(MSE) %>% 
    forestmangr::round_df(8)->All_MSE_AM_3

  All_MSE<-rbind(data.frame("Model"="Real Model",All_MSE_RM),
                 data.frame("Model"="Assumed Model 1",All_MSE_AM_1),
                 data.frame("Model"="Assumed Model 2",All_MSE_AM_2),
                 data.frame("Model"="Assumed Model 3",All_MSE_AM_3))

  remove(All_MSE_RM,All_MSE_AM_1,All_MSE_AM_2,All_MSE_AM_3)

  # Summarise Bias
  Est_Bias_RM %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + 
                                 Theta2*Theta2 + Theta3*Theta3)) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(8)->All_Bias_RM

  Est_Bias_AM_1 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + 
                                 Theta2*Theta2)) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(8)->All_Bias_AM_1

   Est_Bias_AM_2 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + 
                                 Theta2*Theta2+ Theta3*Theta3)) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(8)->All_Bias_AM_2

  Est_Bias_AM_3 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3,Theta4) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + 
                                 Theta2*Theta2+ Theta3*Theta3+ Theta4*Theta4)) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(8)->All_Bias_AM_3

  All_Bias<-rbind(data.frame("Model"="Real Model",All_Bias_RM),
                  data.frame("Model"="Assumed Model 1",All_Bias_AM_1),
                  data.frame("Model"="Assumed Model 2",All_Bias_AM_2),
                  data.frame("Model"="Assumed Model 3",All_Bias_AM_3))

  remove(All_Bias_AM_1,All_Bias_AM_2,All_Bias_AM_3,All_Bias_RM)

  # Summarise Bias = sqrt(Bias^2)
  Est_Bias_RM %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 + 
                          Theta2*Theta2 + Theta3*Theta3 ) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(4)->All_Bias_squared_RM

  Est_Bias_AM_1 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 + 
                            Theta2*Theta2) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(4)->All_Bias_squared_AM_1

  Est_Bias_AM_2 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 + 
                            Theta2*Theta2 + Theta3*Theta3) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(4)->All_Bias_squared_AM_2

  Est_Bias_AM_3 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3,Theta4) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 + 
                            Theta2*Theta2 + Theta3*Theta3 + Theta4*Theta4) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(4)->All_Bias_squared_AM_3

  All_Bias_squared<-rbind(data.frame("Model"="Real Model",All_Bias_squared_RM),
                          data.frame("Model"="Assumed Model 1",All_Bias_squared_AM_1),
                          data.frame("Model"="Assumed Model 2",All_Bias_squared_AM_2),
                          data.frame("Model"="Assumed Model 3",All_Bias_squared_AM_3))

  remove(All_Bias_squared_RM,All_Bias_squared_AM_1,All_Bias_squared_AM_2,All_Bias_squared_AM_3)
}

# Model 4
if(params$Model_Path=="Model_4")
{
  # Summarize SMSE
  Est_Parameter_RM %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3,Theta4) %>%
    group_by(Type,Subsample_Size) %>%
    mutate( MSE=Final_MSE(Real = Real_Vline$Values,
                            Estimated_1=Theta0, 
                            Estimated_2=Theta1,
                            Estimated_3=Theta2,
                            Estimated_4=Theta3,
                            Estimated_5=Theta4,
                            Model = params$Model_Path ))%>%
    distinct(MSE) %>% 
    forestmangr::round_df(8)->All_MSE_RM

  Est_Parameter_AM_1 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2) %>%
    group_by(Type,Subsample_Size) %>%
    mutate( MSE=Final_MSE(Real = Estimated_Vline_AM_1$Values,
                            Estimated_1=Theta0, 
                            Estimated_2=Theta1,
                            Estimated_3=Theta2,
                            Model = "Model_1" ))%>%
    distinct(MSE) %>% 
    forestmangr::round_df(8)->All_MSE_AM_1

  Est_Parameter_AM_2 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate( MSE=Final_MSE(Real = Estimated_Vline_AM_2$Values,
                            Estimated_1=Theta0, 
                            Estimated_2=Theta1,
                            Estimated_3=Theta2,
                            Estimated_4=Theta3,
                            Model = "Model_3" ))%>%
    distinct(MSE) %>% 
    forestmangr::round_df(8)->All_MSE_AM_2

  Est_Parameter_AM_3 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate( MSE=Final_MSE(Real = Estimated_Vline_AM_3$Values,
                            Estimated_1=Theta0, 
                            Estimated_2=Theta1,
                            Estimated_3=Theta2,
                            Estimated_4=Theta3,
                            Model = "Model_2" ))%>%
    distinct(MSE) %>% 
    forestmangr::round_df(8)->All_MSE_AM_3

  All_MSE<-rbind(data.frame("Model"="Real Model",All_MSE_RM),
                 data.frame("Model"="Assumed Model 1",All_MSE_AM_1),
                 data.frame("Model"="Assumed Model 2",All_MSE_AM_2),
                 data.frame("Model"="Assumed Model 3",All_MSE_AM_3))

  remove(All_MSE_RM,All_MSE_AM_1,All_MSE_AM_2,All_MSE_AM_3)

  # Summarise Bias
  Est_Bias_RM %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3,Theta4) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + 
                                 Theta2*Theta2 + Theta3*Theta3 + Theta4*Theta4)) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(8)->All_Bias_RM

  Est_Bias_AM_1 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + 
                                 Theta2*Theta2)) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(8)->All_Bias_AM_1

   Est_Bias_AM_2 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + 
                                 Theta2*Theta2+ Theta3*Theta3)) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(8)->All_Bias_AM_2

  Est_Bias_AM_3 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + 
                                 Theta2*Theta2+ Theta3*Theta3)) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(8)->All_Bias_AM_3

  All_Bias<-rbind(data.frame("Model"="Real Model",All_Bias_RM),
                  data.frame("Model"="Assumed Model 1",All_Bias_AM_1),
                  data.frame("Model"="Assumed Model 2",All_Bias_AM_2),
                  data.frame("Model"="Assumed Model 3",All_Bias_AM_3))

  remove(All_Bias_AM_1,All_Bias_AM_2,All_Bias_AM_3,All_Bias_RM)

  # Summarise Bias = sqrt(Bias^2)
  Est_Bias_RM %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3,Theta4) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 + 
                          Theta2*Theta2 + Theta3*Theta3 + Theta4*Theta4) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(4)->All_Bias_squared_RM

  Est_Bias_AM_1 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 + 
                            Theta2*Theta2) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(4)->All_Bias_squared_AM_1

  Est_Bias_AM_2 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 + 
                            Theta2*Theta2 + Theta3*Theta3) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(4)->All_Bias_squared_AM_2

  Est_Bias_AM_3 %>%
    dplyr::select(Type,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Subsample_Size) %>%
    mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 + 
                            Theta2*Theta2 + Theta3*Theta3 ) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(4)->All_Bias_squared_AM_3

  All_Bias_squared<-rbind(data.frame("Model"="Real Model",All_Bias_squared_RM),
                          data.frame("Model"="Assumed Model 1",All_Bias_squared_AM_1),
                          data.frame("Model"="Assumed Model 2",All_Bias_squared_AM_2),
                          data.frame("Model"="Assumed Model 3",All_Bias_squared_AM_3))

  remove(All_Bias_squared_RM,All_Bias_squared_AM_1,All_Bias_squared_AM_2,All_Bias_squared_AM_3)
}

# Summarise EMSE
inner_join(All_VAR,All_Bias_squared,by=c('Type',"Model","Subsample_Size")) %>%
  mutate(EMSE=mean_Variance+mean_Bias) %>%
  select(Type,Model,Subsample_Size,EMSE)->All_EMSE
colnames(All_EMSE)<-c('Type',"Model","Subsample_Size","EMSE")

if(params$Model_Path=="Model_1")
{
  Full_SMSE_RM<-Final_MSE(Real=Real_Vline$Values,
                          Estimated_1 = Full_Est_Parameter_RM$Theta0,
                          Estimated_2 = Full_Est_Parameter_RM$Theta1,
                          Estimated_3 = Full_Est_Parameter_RM$Theta2,
                          Model=params$Model_Path)
}

if(params$Model_Path=="Model_2" || params$Model_Path=="Model_3")
{
  Full_SMSE_RM<-Final_MSE(Real=Real_Vline$Values,
                          Estimated_1 = Full_Est_Parameter_RM$Theta0,
                          Estimated_2 = Full_Est_Parameter_RM$Theta1,
                          Estimated_3 = Full_Est_Parameter_RM$Theta2,
                          Estimated_4 = Full_Est_Parameter_RM$Theta3,
                          Model=params$Model_Path)
}

if(params$Model_Path=="Model_4")
{
  Full_SMSE_RM<-Final_MSE(Real=Real_Vline$Values,
                          Estimated_1 = Full_Est_Parameter_RM$Theta0,
                          Estimated_2 = Full_Est_Parameter_RM$Theta1,
                          Estimated_3 = Full_Est_Parameter_RM$Theta2,
                          Estimated_4 = Full_Est_Parameter_RM$Theta3,
                          Estimated_5 = Full_Est_Parameter_RM$Theta4,
                          Model=params$Model_Path)
}

Full_EMSE_RM<-Full_Variance_RM+Full_Bias_RM*Full_Bias_RM
Full_EMSE_AM_1<-Full_Variance_AM_1+Full_Bias_AM_1*Full_Bias_AM_1
Full_EMSE_AM_2<-Full_Variance_AM_2+Full_Bias_AM_2*Full_Bias_AM_2
Full_EMSE_AM_3<-Full_Variance_AM_3+Full_Bias_AM_3*Full_Bias_AM_3

remove(Temp_P,Temp_W,Temp_Info)

# Store the Results
save(All_MSE,All_EMSE,All_Bias,All_VAR,All_Info,
     file = here("Non_Identical_r0","Summary","Classical",params$Model_Path,"Data","Efficiency",
                 "No_Correlated_Covariate_OSMAC_MF.RData"))

```

## SMSE, EMSE, A-optimality, D-optimality, Bias Comparing Real Model with with Full Data Model

```{r Final plots 1,fig.width=20,fig.height=8}
Plot_list<-list()

All_MSE[All_MSE$Model=="Real Model",] %>%
  #ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=MSE/Full_SMSE_RM) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+facet_wrap(~Type,ncol = 1)+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("SMSE Efficiency")->Plot_list[[1]]

All_VAR[All_VAR$Model=="Real Model",] %>%
  ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Variance/Full_Variance_RM) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+facet_wrap(~Type,ncol = 1)+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("Estimated Variance Efficiency")->Plot_list[[2]]

All_Info[All_Info$Model=="Real Model",] %>%
  ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Info/Full_Information_RM) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+facet_wrap(~Type,ncol = 1)+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("Estimated Model Information Efficiency")->Plot_list[[3]]

All_Bias[All_Bias$Model=="Real Model",] %>%
  #ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Bias/Full_Bias_RM) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+facet_wrap(~Type,ncol = 1)+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("Estimated Bias Efficiency")->Plot_list[[4]]

All_EMSE[All_EMSE$Model=="Real Model",] %>%
  #ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=EMSE/Full_EMSE_RM) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+facet_wrap(~Type,ncol = 1)+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("EMSE Efficiency")->Plot_list[[5]]

ggarrange(plotlist = Plot_list,nrow = 1,ncol = 5,common.legend = TRUE)

```

## EMSE, A-optimality, D-optimality, Bias Comparing Assumed Model with Full Data Assumed Model 1

```{r Final plots AM 1,fig.width=16,fig.height=8}
Plot_list<-list()

All_VAR[All_VAR$Model=="Assumed Model 1",] %>%
  ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Variance/Full_Variance_AM_1) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+facet_wrap(~Type,ncol = 1)+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("Estimated Variance Efficiency")->Plot_list[[1]]

All_Info[All_Info$Model=="Assumed Model 1",] %>%
  ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Info/Full_Information_AM_1) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+facet_wrap(~Type,ncol = 1)+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("Estimated Model Information Efficiency")->Plot_list[[2]]

All_Bias[All_Bias$Model=="Assumed Model 1",] %>%
  #ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Bias/Full_Bias_AM_1) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+facet_wrap(~Type,ncol = 1)+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("Estimated Bias Efficiency")->Plot_list[[3]]

All_EMSE[All_EMSE$Model=="Assumed Model 1",] %>%
  #ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=EMSE/Full_EMSE_AM_1) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+facet_wrap(~Type,ncol = 1)+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("EMSE Efficiency")->Plot_list[[4]]

ggarrange(plotlist = Plot_list,nrow = 1,ncol = 4,common.legend = TRUE)
```

## EMSE, A-optimality, D-optimality, Bias Comparing Assumed Model with Full Data Assumed Model 2

```{r Final plots AM 2,fig.width=16,fig.height=8}
Plot_list<-list()

All_VAR[All_VAR$Model=="Assumed Model 2",] %>%
  ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Variance/Full_Variance_AM_2) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+facet_wrap(~Type,ncol = 1)+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("Estimated Variance Efficiency")->Plot_list[[1]]

All_Info[All_Info$Model=="Assumed Model 2",] %>%
  ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Info/Full_Information_AM_2) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+facet_wrap(~Type,ncol = 1)+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("Estimated Model Information Efficiency")->Plot_list[[2]]

All_Bias[All_Bias$Model=="Assumed Model 2",] %>%
  #ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Bias/Full_Bias_AM_2) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+facet_wrap(~Type,ncol = 1)+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("Estimated Bias Efficiency")->Plot_list[[3]]

All_EMSE[All_EMSE$Model=="Assumed Model 2",] %>%
  #ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=EMSE/Full_EMSE_AM_2) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+facet_wrap(~Type,ncol = 1)+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("EMSE Efficiency")->Plot_list[[4]]

ggarrange(plotlist = Plot_list,nrow = 1,ncol = 4,common.legend = TRUE)
```

## EMSE, A-optimality, D-optimality, Bias Comparing Assumed Model with Full Data Assumed Model 3

```{r Final plots AM 3,fig.width=16,fig.height=8}
Plot_list<-list()

All_VAR[All_VAR$Model=="Assumed Model 3",] %>%
  ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Variance/Full_Variance_AM_3) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+facet_wrap(~Type,ncol = 1)+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("Estimated Variance Efficiency")->Plot_list[[1]]

All_Info[All_Info$Model=="Assumed Model 3",] %>%
  ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Info/Full_Information_AM_3) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+facet_wrap(~Type,ncol = 1)+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("Estimated Model Information Efficiency")->Plot_list[[2]]

All_Bias[All_Bias$Model=="Assumed Model 3",] %>%
  #ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Bias/Full_Bias_AM_3) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+facet_wrap(~Type,ncol = 1)+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("Estimated Bias Efficiency")->Plot_list[[3]]

All_EMSE[All_EMSE$Model=="Assumed Model 3",] %>%
  #ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=EMSE/Full_EMSE_AM_3) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+facet_wrap(~Type,ncol = 1)+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("EMSE Efficiency")->Plot_list[[4]]

ggarrange(plotlist = Plot_list,nrow = 1,ncol = 4,common.legend = TRUE)
```