### 2022-01-17 Non_Identical_r0/Rmarkdown/Classical/Random_Sampling.Rmd

---
title: "Random_Sampling; Model: `r params$Model_Path`;"
author: "M.Amalan"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,fig.height = 9,fig.width = 9,message = FALSE,warning = FALSE)
library(here)
library(ggplot2)
library(dplyr)
library(tidyr)
library(LaplacesDemon)
library(linguisticsdown)
library(gganimate)
library(viridis)
library(compiler)
library(ggpubr)

# Cordeiro Bias Estimation ----
Cordeiro<-function(XData,With_bias)
{
  p <- as.vector(exp(XData%*%as.vector(With_bias)))
  W <- diag(p)
  inverse_term <- solve(t(XData)%*%W%*%XData)

  Term1 <- inverse_term%*%t(XData)%*%W
  Term2 <- diag(diag(XData%*%(inverse_term)%*%t(XData))) %*% rep(-0.5,nrow(XData))

  bias <- as.vector(Term1%*%Term2)
  return(bias)
}

Cordeiro<-cmpfun(Cordeiro)

FullDataPath<-here("Non_Identical_r0","Generate_Big_Data",params$Model_Path,"No_Correlated_Covariate.RData")
load(here(FullDataPath))
Full_model<-glm(Y~.-1,data = as.data.frame(Simulated_Data[[2]]$Real_Model),family = "poisson")
Real_Vline<-data.frame("Theta"=c(paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1]))),
                       "Values"=Simulated_Data$Basic$Theta)
Estimated_Vline<-data.frame("Theta"=c(paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1]))),
                            "Values"=Full_model$coefficients)
Bias_Vline<-data.frame("Theta"=c(paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1]))),"Values"=0)
```

# Random Sampling; Model : `r params$Model_Path`

# Real Model Data

```{r Load Data RM}
Path<-here("Non_Identical_r0","Outputs","Classical",params$Model_Path,"Random_Sampling","Real_Model/")
load(here(Path,"Random_Sample_output.RData"))

Est_Parameter_RM<-as.data.frame(Est_Param_RandomSample)
colnames(Est_Parameter_RM)<-c("Subsample_Size",paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))

Utility_RM<-as.data.frame(Utility_RandomSample) 
colnames(Utility_RM)<-c("Subsample_Size",paste0(c("A","D"),"_optimality"))

Est_Bias_RM<-as.data.frame(Bias_RandomSample) 
colnames(Est_Bias_RM)<-c("Subsample_Size",paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))

remove(Est_Param_RandomSample,Utility_RandomSample,Bias_RandomSample#,SelectedData_RandomSample
       )
```

# Assumed Model 1 Data

```{r Load Data AM 1}
Path<-here("Non_Identical_r0","Outputs","Classical",params$Model_Path,"Random_Sampling","Assumed_Model_1/")
load(here(Path,"Random_Sample_output.RData"))

Est_Parameter_AM_1<-as.data.frame(Est_Param_RandomSample)
colnames(Est_Parameter_AM_1)<-c("Subsample_Size",paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))

Utility_AM_1<-as.data.frame(Utility_RandomSample) 
colnames(Utility_AM_1)<-c("Subsample_Size",paste0(c("A","D"),"_optimality"))

Est_Bias_AM_1<-as.data.frame(Bias_RandomSample) 
colnames(Est_Bias_AM_1)<-c("Subsample_Size",paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))

remove(Est_Param_RandomSample,Utility_RandomSample,Bias_RandomSample#,SelectedData_RandomSample
       )
```

# Assumed Model 2 Data

```{r Load Data AM 2}
Path<-here("Non_Identical_r0","Outputs","Classical",params$Model_Path,"Random_Sampling","Assumed_Model_2/")
load(here(Path,"Random_Sample_output.RData"))

Est_Parameter_AM_2<-as.data.frame(Est_Param_RandomSample)
colnames(Est_Parameter_AM_2)<-c("Subsample_Size",paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))

Utility_AM_2<-as.data.frame(Utility_RandomSample) 
colnames(Utility_AM_2)<-c("Subsample_Size",paste0(c("A","D"),"_optimality"))

Est_Bias_AM_2<-as.data.frame(Bias_RandomSample) 
colnames(Est_Bias_AM_2)<-c("Subsample_Size",paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))

remove(Est_Param_RandomSample,Utility_RandomSample,Bias_RandomSample#,SelectedData_RandomSample
       )
```

# Assumed Model 3 Data

```{r Load Data AM 3}
Path<-here("Non_Identical_r0","Outputs","Classical",params$Model_Path,"Random_Sampling","Assumed_Model_3/")
load(here(Path,"Random_Sample_output.RData"))

Est_Parameter_AM_3<-as.data.frame(Est_Param_RandomSample)
colnames(Est_Parameter_AM_3)<-c("Subsample_Size",paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))

Utility_AM_3<-as.data.frame(Utility_RandomSample) 
colnames(Utility_AM_3)<-c("Subsample_Size",paste0(c("A","D"),"_optimality"))

Est_Bias_AM_3<-as.data.frame(Bias_RandomSample) 
colnames(Est_Bias_AM_3)<-c("Subsample_Size",paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))

remove(Est_Param_RandomSample,Utility_RandomSample,Bias_RandomSample#,SelectedData_RandomSample
       )
```

```{r Combine Datasets}
# Combine datasets
Est_Parameter<-rbind(data.frame("Model"="Real Model",Est_Parameter_RM),
                     data.frame("Model"="Assumed Model 1",Est_Parameter_AM_1),
                     data.frame("Model"="Assumed Model 2",Est_Parameter_AM_2),
                     data.frame("Model"="Assumed Model 3",Est_Parameter_AM_3))

Est_Bias<-rbind(data.frame("Model"="Real Model",Est_Bias_RM),
                data.frame("Model"="Assumed Model 1",Est_Bias_AM_1),
                data.frame("Model"="Assumed Model 2",Est_Bias_AM_2),
                data.frame("Model"="Assumed Model 3",Est_Bias_AM_3))

Est_Utility<-rbind(data.frame("Model"="Real Model",Utility_RM),
                   data.frame("Model"="Assumed Model 1",Utility_AM_1),
                   data.frame("Model"="Assumed Model 2",Utility_AM_2),
                   data.frame("Model"="Assumed Model 3",Utility_AM_3))

remove(Est_Bias_AM_1,Est_Bias_AM_2,Est_Bias_AM_3,Est_Bias_RM,
       Est_Parameter_AM_1,Est_Parameter_AM_2,Est_Parameter_AM_3,Est_Parameter_RM,
       Utility_AM_1,Utility_AM_2,Utility_AM_3,Utility_RM)
```

## Parameters

```{r Animated Plots}
Est_Parameter %>%
  pivot_longer(starts_with("Theta"), names_to = "Theta", values_to = "Estimated") %>%
ggplot(., aes(x = Estimated, fill=Model, frame=Subsample_Size)) + 
  geom_histogram(alpha=0.8)+
  facet_grid(~Theta,scales = "free_x",shrink = TRUE)+
  geom_vline(data= Estimated_Vline,aes(xintercept = Values,color="red"),linetype='dashed',size=1)+
  scale_color_discrete(name = "Parameter values", labels = c("Estimated"))+
  labs(title=paste0("Real Model :",params$Model_Path," Data : Estimated Theta"),
      subtitle = "Sub-sample Size: {current_frame}")+
  transition_manual(Subsample_Size)+
  theme_light()+ylab("Frequency")+xlab("Estimates")->para_1

anim_save(filename="para_1.gif",animation = para_1,
          path=here("Non_Identical_r0","htmloutputs","Classical",params$Model_Path,"Random_Sampling","Plots"))

include_graphics2(here("Non_Identical_r0","htmloutputs","Classical",params$Model_Path,"Random_Sampling","Plots","para_1.gif"))
```

## Bias

```{r Animated Plots for Bias}
Est_Bias %>%
  pivot_longer(starts_with("Theta"), names_to = "Theta", values_to = "Estimated") %>%
ggplot(., aes(x = Estimated, fill=Model, frame=Subsample_Size)) + 
  geom_histogram(alpha=0.8)+
  facet_grid(~Theta,scales = "free_x",shrink = TRUE)+
  geom_vline(data= Bias_Vline,aes(xintercept = Values,color="red"),linetype='dashed',size=1)+
  scale_color_discrete(name = "Bias values", labels = c("Real"))+
  labs(title=paste0("Real Model :",params$Model_Path," Data : Estimated Bias"),
      subtitle = "Sub-sample Size: {current_frame}")+
  transition_manual(Subsample_Size)+theme_light()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  ylab("Frequency")+xlab("Estimates")->Bias_1

anim_save(filename="Bias_1.gif",animation = Bias_1,
          path=here("Non_Identical_r0","htmloutputs","Classical",params$Model_Path,"Random_Sampling","Plots"))

include_graphics2(here("Non_Identical_r0","htmloutputs","Classical",params$Model_Path,"Random_Sampling","Plots","Bias_1.gif"))
```

## Utility values

```{r Animated Utility plots RM,fig.width=7,fig.height=7}
Est_Utility %>%
  ggplot(.,aes(x=log(A_optimality),fill=Model))+
  geom_histogram()+
  theme_light()+ylab("Frequency")+xlab("log(A-optimality)")+
  transition_states(Subsample_Size)+
  ggtitle("Real Model : Subsample : {closest_state}")->U_1

anim_save(filename="U_1.gif",animation = U_1,
          path=here("Non_Identical_r0","htmloutputs","Classical",params$Model_Path,"Random_Sampling","Plots"))

include_graphics2(here("Non_Identical_r0","htmloutputs","Classical",params$Model_Path,"Random_Sampling","Plots","U_1.gif"))

Est_Utility %>%
  ggplot(.,aes(x=log(D_optimality),fill=Model))+
  geom_histogram()+
  theme_light()+ylab("Frequency")+xlab("log(D-optimality) ")+
  transition_states(Subsample_Size)+
  ggtitle("Real Model : Subsample : {closest_state}")->U_2

anim_save(filename="U_2.gif",animation = U_2,
          path=here("Non_Identical_r0","htmloutputs","Classical",params$Model_Path,"Random_Sampling","Plots"))

include_graphics2(here("Non_Identical_r0","htmloutputs","Classical",params$Model_Path,"Random_Sampling","Plots","U_2.gif"))
```

## Response Data of the Subsamples

```{r Response Data RM,eval=FALSE}
Random_Sample %>%
  as.data.frame()%>%
  group_by(Simulation) %>%
  dplyr::count(Y)-> Response

remove_missing(Response) %>% 
  ggplot(.,aes(x=Simulation,y=n,fill=as.factor(Y)))+
  geom_bar(stat = "identity")+
  scale_fill_viridis_d()+theme_light()+
  ylab("Count / Frequency")+xlab("Simulation")+
  labs(fill=c("Response Variable"))+
  ggtitle("Real Model : Dispersion of Y variable after multiple simulations")
```

# Comparison for Efficiency

```{r Comparison of Summary Data}
Final_MSE<-function(Real,Estimated_1,Estimated_2,Estimated_3,Estimated_4,Estimated_5,Model)
{
  For_Theta0<-sum((Real[1]-Estimated_1)^2,na.rm = TRUE)
  For_Theta1<-sum((Real[2]-Estimated_2)^2,na.rm = TRUE)
  For_Theta2<-sum((Real[3]-Estimated_3)^2,na.rm = TRUE)

  if(Model=="Model_1")
  {
    return((For_Theta0+For_Theta1+For_Theta2)/length(Estimated_1))
  }  
  if(Model=="Model_2" || Model=="Model_3")
  {
    For_Theta3<-sum((Real[4]-Estimated_4)^2,na.rm = TRUE)
    return((For_Theta0+For_Theta1+For_Theta2+For_Theta3)/length(Estimated_1))
  }
  if(Model=="Model_4")
  {
    For_Theta3<-sum((Real[4]-Estimated_4)^2,na.rm = TRUE)
    For_Theta4<-sum((Real[5]-Estimated_5)^2,na.rm = TRUE)
    return((For_Theta0+For_Theta1+For_Theta2+For_Theta3+For_Theta4)/length(Estimated_1))
  }
}

# Summarise A- optimality
Est_Utility %>%
  dplyr::select(Model,Subsample_Size,A_optimality) %>%
  group_by(Model,Subsample_Size) %>%
  mutate(mean_Variance=mean(A_optimality))%>%
  distinct(mean_Variance) %>% 
  forestmangr::round_df(8)->All_VAR

# Summarise D- optimality
Est_Utility %>%
  dplyr::select(Model,Subsample_Size,D_optimality) %>%
  group_by(Model,Subsample_Size) %>%
  mutate(mean_Info=mean(D_optimality))%>%
  distinct(mean_Info) %>% 
  forestmangr::round_df(8)->All_Info

if(params$Model_Path=="Model_1")
{
  # Summarize SMSE
  Est_Parameter %>%
        dplyr::select(Model,Subsample_Size,Theta0,Theta1,Theta2) %>%
        group_by(Model,Subsample_Size) %>%
        mutate( MSE=Final_MSE(Real = Real_Vline$Values,
                              Estimated_1=Theta0, 
                              Estimated_2=Theta1,
                              Estimated_3=Theta2,
                              Model = params$Model_Path ))%>%
        distinct(MSE) %>% 
        forestmangr::round_df(8)->All_MSE

# Summarise Bias
Est_Bias %>%
  dplyr::select(Model,Subsample_Size,Theta0,Theta1,Theta2) %>%
  group_by(Model,Subsample_Size) %>%
  mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + Theta2*Theta2)) )%>%
  distinct(mean_Bias) %>% 
  forestmangr::round_df(8)->All_Bias

# Summarise Bias = sqrt(Bias^2)
Est_Bias %>%
  dplyr::select(Model,Subsample_Size,Theta0,Theta1,Theta2) %>%
  group_by(Model,Subsample_Size) %>%
  mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 + Theta2*Theta2) )%>%
  distinct(mean_Bias) %>% 
  forestmangr::round_df(4)->All_Bias_squared
}

if(params$Model_Path=="Model_2"||params$Model_Path=="Model_3")
{
  # Summarize SMSE
  Est_Parameter %>%
        dplyr::select(Model,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
        group_by(Model,Subsample_Size) %>%
        mutate( MSE=Final_MSE(Real = Real_Vline$Values,
                              Estimated_1=Theta0, 
                              Estimated_2=Theta1,
                              Estimated_3=Theta2,
                              Estimated_4=Theta3,
                              Model = params$Model_Path ))%>%
        distinct(MSE) %>% 
        forestmangr::round_df(8)->All_MSE

# Summarise Bias
Est_Bias %>%
  dplyr::select(Model,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
  group_by(Model,Subsample_Size) %>%
  mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + 
                             Theta2*Theta2 + Theta3*Theta3)) )%>%
  distinct(mean_Bias) %>% 
  forestmangr::round_df(8)->All_Bias

# Summarise Bias = sqrt(Bias^2)
Est_Bias %>%
  dplyr::select(Model,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
  group_by(Model,Subsample_Size) %>%
  mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 + 
                        Theta2*Theta2 + Theta3*Theta3 ) )%>%
  distinct(mean_Bias) %>% 
  forestmangr::round_df(4)->All_Bias_squared
}

if(params$Model_Path=="Model_4")
{
  # Summarize SMSE
  Est_Parameter %>%
        dplyr::select(Model,Subsample_Size,Theta0,Theta1,Theta2,Theta3,Theta4) %>%
        group_by(Model,Subsample_Size) %>%
        mutate( MSE=Final_MSE(Real = Real_Vline$Values,
                              Estimated_1=Theta0, 
                              Estimated_2=Theta1,
                              Estimated_3=Theta2,
                              Estimated_4=Theta3,
                              Estimated_5=Theta4,
                              Model = params$Model_Path ))%>%
        distinct(MSE) %>% 
        forestmangr::round_df(8)->All_MSE

# Summarise Bias
Est_Bias %>%
  dplyr::select(Model,Subsample_Size,Theta0,Theta1,Theta2,Theta3,Theta4) %>%
  group_by(Model,Subsample_Size) %>%
  mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + 
                             Theta2*Theta2 + Theta3*Theta3 + Theta4*Theta4)) )%>%
  distinct(mean_Bias) %>% 
  forestmangr::round_df(8)->All_Bias

# Summarise Bias = sqrt(Bias^2)
Est_Bias %>%
  dplyr::select(Model,Subsample_Size,Theta0,Theta1,Theta2,Theta3,Theta4) %>%
  group_by(Model,Subsample_Size) %>%
  mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 + 
                        Theta2*Theta2 + Theta3*Theta3 + Theta4*Theta4) )%>%
  distinct(mean_Bias) %>% 
  forestmangr::round_df(4)->All_Bias_squared
}

# Summarise EMSE
inner_join(All_VAR,All_Bias_squared,by=c("Model","Subsample_Size")) %>%
  mutate(EMSE=mean_Variance+mean_Bias) %>%
  select(Model,Subsample_Size,EMSE)->All_EMSE
colnames(All_EMSE)<-c("Model","Subsample_Size","EMSE")

# Full Data model Estimates
Full_Est_Parameter<-as.data.frame(t(Estimated_Vline$Values))
Full_Est_Bias<-as.data.frame(t(Cordeiro(XData = Simulated_Data[[2]]$Real_Model[,-1],
                                      With_bias = as.matrix(Full_Est_Parameter))))
colnames(Full_Est_Parameter)<-colnames(Full_Est_Bias)<-c(paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))

Full_Bias<-sqrt(sum(Full_Est_Bias^2))
Temp_P<-exp(Simulated_Data[[2]]$Real_Model[,-1]%*% t(as.matrix(Full_Est_Parameter)))
Temp_W<-diag(as.vector(Temp_P))
Temp_Info<-t(Simulated_Data[[2]]$Real_Model[,-1])%*% Temp_W %*% Simulated_Data[[2]]$Real_Model[,-1]
Full_Variance<-tr(solve(Temp_Info))
Full_Information<-det(Temp_Info)

if(params$Model_Path=="Model_1")
{
  Full_SMSE<-Final_MSE(Real=Real_Vline$Values,
                Estimated_1 = Full_Est_Parameter$Theta0,
                Estimated_2 = Full_Est_Parameter$Theta1,
                Estimated_3 = Full_Est_Parameter$Theta2,
                Model=params$Model_Path)
}

if(params$Model_Path=="Model_2"||params$Model_Path=="Model_3")
{
  Full_SMSE<-Final_MSE(Real=Real_Vline$Values,
                Estimated_1 = Full_Est_Parameter$Theta0,
                Estimated_2 = Full_Est_Parameter$Theta1,
                Estimated_3 = Full_Est_Parameter$Theta2,
                Estimated_4 = Full_Est_Parameter$Theta3,
                Model=params$Model_Path)
}

if(params$Model_Path=="Model_4")
{
  Full_SMSE<-Final_MSE(Real=Real_Vline$Values,
                Estimated_1 = Full_Est_Parameter$Theta0,
                Estimated_2 = Full_Est_Parameter$Theta1,
                Estimated_3 = Full_Est_Parameter$Theta2,
                Estimated_4 = Full_Est_Parameter$Theta3,
                Estimated_5 = Full_Est_Parameter$Theta4,
                Model=params$Model_Path)
}

Full_EMSE<-Full_Variance+Full_Bias*Full_Bias

remove(Temp_P,Temp_W,Temp_Info)

# Store the Results
save(All_MSE,All_EMSE,All_Bias,All_VAR,All_Info,
     file = here("Non_Identical_r0","Summary","Classical",params$Model_Path,"Data","Efficiency",
                 "No_Correlated_Covariate_RS.RData"))
```

## SMSE, EMSE, A-optimality, D-optimality, Bias Comparing Real Model and Assumed Models

```{r Final plots 1,fig.width=6,fig.height=20}
Plot_list<-list()

All_MSE %>%
  ungroup(Model,Subsample_Size) %>%
ggplot(.,aes(x=Subsample_Size,y=MSE,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("SMSE")+ggtitle("SMSE Comparison")->Plot_list[[1]]

All_VAR %>%
  ungroup(Model,Subsample_Size) %>%
ggplot(.,aes(x=Subsample_Size,y=mean_Variance,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("Mean Variance")+ggtitle("Estimated Variance Comparison")->Plot_list[[2]]

All_Info %>%
  ungroup(Model,Subsample_Size) %>%
ggplot(.,aes(x=Subsample_Size,y=mean_Info,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("Average Information")+ggtitle("Estimated Model Information Comparison")->Plot_list[[3]]

All_Bias %>%
  ungroup(Model,Subsample_Size) %>%
ggplot(.,aes(x=Subsample_Size,y=mean_Bias,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("Average Bias")+ggtitle("Estimated Bias Comparison")->Plot_list[[4]]

All_EMSE %>%
  ungroup(Model,Subsample_Size) %>%
ggplot(.,aes(x=Subsample_Size,y=EMSE,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("EMSE")+ggtitle("EMSE Comparison")->Plot_list[[5]]

ggarrange(plotlist = Plot_list,nrow = 5,ncol = 1,common.legend = TRUE)

```

## SMSE, EMSE, A-optimality, D-optimality, Bias Comparing Real Model and Assumed Models with Full Data Model

```{r Final plots 2,fig.width=6,fig.height=20}
Plot_list<-list()
All_MSE %>%
  ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=MSE/Full_SMSE) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("SMSE Efficiency")->Plot_list[[1]]

All_VAR %>%
  ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Variance/Full_Variance) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("Estimated Variance Efficiency")->Plot_list[[2]]

All_Info %>%
  ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Info/Full_Information) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("Estimated Model Information Efficiency")->Plot_list[[3]]

All_Bias %>%
  ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Bias/Full_Bias) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("Estimated Bias Efficiency")->Plot_list[[4]]

All_EMSE %>%
  ungroup(Model,Subsample_Size) %>%
  mutate(Efficiency=EMSE/Full_EMSE) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("Relative Efficiency")+ggtitle("EMSE Efficiency")->Plot_list[[5]]

ggarrange(plotlist = Plot_list,nrow = 5,ncol = 1,common.legend = TRUE)
```