### 2022-01-17 Non_Identical_r0/Rmarkdown/Classical/Summary/Best_Subsampling_Method.Rmd

---
title: "Best Sub-sampling Method: `r params$Model_Path`"
author: "M.Amalan"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,fig.width = 12,fig.height = 12,message = FALSE,warning = FALSE)
library(here)
library(ggplot2)
library(dplyr)
library(tidyr)
library(LaplacesDemon)
library(xtable)
library(flextable)
library(grid)
library(cowplot)
library(linguisticsdown)
library(ggpubr)
```

```{r load Data}
Old_Labels<-c("Real Model",paste0("Assumed Model ",1:3))
New_Labels<-c("RM",paste0("AM ",1:3))

# Random Sampling
load(here("Non_Identical_r0","Summary","Classical",params$Model_Path,"Data","Efficiency",
          "No_Correlated_Covariate_RS.RData"))

for (i in 1:length(Old_Labels)) 
{
  All_Bias$Model[All_Bias$Model==Old_Labels[i]] <-paste0("RS ",New_Labels[i])
  All_EMSE$Model[All_EMSE$Model==Old_Labels[i]] <-paste0("RS ",New_Labels[i])
  All_Info$Model[All_Info$Model==Old_Labels[i]] <-paste0("RS ",New_Labels[i])
  All_MSE$Model[All_MSE$Model==Old_Labels[i]] <-paste0("RS ",New_Labels[i])
  All_VAR$Model[All_VAR$Model==Old_Labels[i]] <-paste0("RS ",New_Labels[i])
}

RS_Bias<-All_Bias 
RS_EMSE<-All_EMSE
RS_Info<-All_Info
RS_MSE<-All_MSE
RS_VAR<-All_VAR

remove(All_Bias,All_EMSE,All_MSE,All_Info,All_VAR)

# OSMAC
load(here("Non_Identical_r0","Summary","Classical",params$Model_Path,"Data","Efficiency",
          "No_Correlated_Covariate_OSMAC.RData"))

## mMSE
for (i in 1:length(Old_Labels)) 
{
  All_Bias$Model[All_Bias$Model==Old_Labels[i] & All_Bias$Type == "mMSE"] <-paste0("OSMAC mMSE ",New_Labels[i])
  All_EMSE$Model[All_EMSE$Model==Old_Labels[i] & All_Bias$Type == "mMSE"] <-paste0("OSMAC mMSE ",New_Labels[i])
  All_Info$Model[All_Info$Model==Old_Labels[i] & All_Bias$Type == "mMSE"] <-paste0("OSMAC mMSE ",New_Labels[i])
  All_MSE$Model[All_MSE$Model==Old_Labels[i] & All_Bias$Type == "mMSE"] <-paste0("OSMAC mMSE ",New_Labels[i])
  All_VAR$Model[All_VAR$Model==Old_Labels[i] & All_Bias$Type == "mMSE"] <-paste0("OSMAC mMSE ",New_Labels[i])
}

OSMAC_mMSE_Bias<-All_Bias[All_Bias$Type == "mMSE",] 
OSMAC_mMSE_EMSE<-All_EMSE[All_EMSE$Type == "mMSE",]
OSMAC_mMSE_Info<-All_Info[All_Info$Type == "mMSE",]
OSMAC_mMSE_MSE<-All_MSE[All_MSE$Type == "mMSE",]
OSMAC_mMSE_VAR<-All_VAR[All_VAR$Type == "mMSE",]

## mVc
for (i in 1:length(Old_Labels)) 
{
  All_Bias$Model[All_Bias$Model==Old_Labels[i] & All_Bias$Type == "mVc"] <-paste0("OSMAC mVc ",New_Labels[i])
  All_EMSE$Model[All_EMSE$Model==Old_Labels[i] & All_Bias$Type == "mVc"] <-paste0("OSMAC mVc ",New_Labels[i])
  All_Info$Model[All_Info$Model==Old_Labels[i] & All_Bias$Type == "mVc"] <-paste0("OSMAC mVc ",New_Labels[i])
  All_MSE$Model[All_MSE$Model==Old_Labels[i] & All_Bias$Type == "mVc"] <-paste0("OSMAC mVc ",New_Labels[i])
  All_VAR$Model[All_VAR$Model==Old_Labels[i] & All_Bias$Type == "mVc"] <-paste0("OSMAC mVc ",New_Labels[i])
}

OSMAC_mVc_Bias<-All_Bias[All_Bias$Type == "mVc",] 
OSMAC_mVc_EMSE<-All_EMSE[All_EMSE$Type == "mVc",]
OSMAC_mVc_Info<-All_Info[All_Info$Type == "mVc",]
OSMAC_mVc_MSE<-All_MSE[All_MSE$Type == "mVc",]
OSMAC_mVc_VAR<-All_VAR[All_VAR$Type == "mVc",]

remove(All_Bias,All_EMSE,All_MSE,All_Info,All_VAR)

# OSMAC Model Free
load(here("Non_Identical_r0","Summary","Classical",params$Model_Path,"Data","Efficiency",
          "No_Correlated_Covariate_OSMAC_MF.RData"))

All_Bias<-All_Bias[All_Bias$Model==Old_Labels[1],]
All_EMSE<-All_EMSE[All_EMSE$Model==Old_Labels[1],]
All_Info<-All_Info[All_Info$Model==Old_Labels[1],]
All_MSE<-All_MSE[All_MSE$Model==Old_Labels[1],]
All_VAR<-All_VAR[All_VAR$Model==Old_Labels[1],]

## mMSE
All_Bias$Model[All_Bias$Type == "mMSE"] <-paste0("OSMAC mMSE MF")
All_EMSE$Model[All_EMSE$Type == "mMSE"] <-paste0("OSMAC mMSE MF")
All_Info$Model[All_Info$Type == "mMSE"] <-paste0("OSMAC mMSE MF")
All_MSE$Model[All_MSE$Type == "mMSE"] <-paste0("OSMAC mMSE MF")
All_VAR$Model[All_VAR$Type == "mMSE"] <-paste0("OSMAC mMSE MF")

OSMAC_MF_mMSE_Bias<-All_Bias[All_Bias$Type == "mMSE",] 
OSMAC_MF_mMSE_EMSE<-All_EMSE[All_EMSE$Type == "mMSE",]
OSMAC_MF_mMSE_Info<-All_Info[All_Info$Type == "mMSE",]
OSMAC_MF_mMSE_MSE<-All_MSE[All_MSE$Type == "mMSE",]
OSMAC_MF_mMSE_VAR<-All_VAR[All_VAR$Type == "mMSE",]

## mVc
All_Bias$Model[All_Bias$Type == "mVc"] <-paste0("OSMAC mVc MF")
All_EMSE$Model[All_EMSE$Type == "mVc"] <-paste0("OSMAC mVc MF")
All_Info$Model[All_Info$Type == "mVc"] <-paste0("OSMAC mVc MF")
All_MSE$Model[All_MSE$Type == "mVc"] <-paste0("OSMAC mVc MF")
All_VAR$Model[All_VAR$Type == "mVc"] <-paste0("OSMAC mVc MF")

OSMAC_MF_mVc_Bias<-All_Bias[All_Bias$Type == "mVc",] 
OSMAC_MF_mVc_EMSE<-All_EMSE[All_EMSE$Type == "mVc",]
OSMAC_MF_mVc_Info<-All_Info[All_Info$Type == "mVc",]
OSMAC_MF_mVc_MSE<-All_MSE[All_MSE$Type == "mVc",]
OSMAC_MF_mVc_VAR<-All_VAR[All_VAR$Type == "mVc",]

remove(All_Bias,All_EMSE,All_MSE,All_Info,All_VAR)

# All elements
All_Bias<-rbind(RS_Bias,
                OSMAC_mMSE_Bias[,!(colnames(OSMAC_mMSE_Bias) %in% "Type")],
                OSMAC_mVc_Bias[,!(colnames(OSMAC_mVc_Bias) %in% "Type")],
                OSMAC_MF_mMSE_Bias[,!(colnames(OSMAC_MF_mMSE_Bias) %in% "Type")],
                OSMAC_MF_mVc_Bias[,!(colnames(OSMAC_MF_mVc_Bias) %in% "Type")])
All_EMSE<-rbind(RS_EMSE,
                OSMAC_mMSE_EMSE[,!(colnames(OSMAC_mMSE_EMSE) %in% "Type")],
                OSMAC_mVc_EMSE[,!(colnames(OSMAC_mVc_EMSE) %in% "Type")],
                OSMAC_MF_mMSE_EMSE[,!(colnames(OSMAC_MF_mMSE_EMSE) %in% "Type")],
                OSMAC_MF_mVc_EMSE[,!(colnames(OSMAC_MF_mVc_EMSE) %in% "Type")])
All_MSE<-rbind(RS_MSE,
                OSMAC_mMSE_MSE[,!(colnames(OSMAC_mMSE_MSE) %in% "Type")],
                OSMAC_mVc_MSE[,!(colnames(OSMAC_mVc_MSE) %in% "Type")],
                OSMAC_MF_mMSE_MSE[,!(colnames(OSMAC_MF_mMSE_MSE) %in% "Type")],
                OSMAC_MF_mVc_MSE[,!(colnames(OSMAC_MF_mVc_MSE) %in% "Type")])
All_Info<-rbind(RS_Info,
                OSMAC_mMSE_Info[,!(colnames(OSMAC_mMSE_Info) %in% "Type")],
                OSMAC_mVc_Info[,!(colnames(OSMAC_mVc_Info) %in% "Type")],
                OSMAC_MF_mMSE_Info[,!(colnames(OSMAC_MF_mMSE_Info) %in% "Type")],
                OSMAC_MF_mVc_Info[,!(colnames(OSMAC_MF_mVc_Info) %in% "Type")])
All_VAR<-rbind(RS_VAR,
                OSMAC_mMSE_VAR[,!(colnames(OSMAC_mMSE_VAR) %in% "Type")],
                OSMAC_mVc_VAR[,!(colnames(OSMAC_mVc_VAR) %in% "Type")],
                OSMAC_MF_mMSE_VAR[,!(colnames(OSMAC_MF_mMSE_VAR) %in% "Type")],
                OSMAC_MF_mVc_VAR[,!(colnames(OSMAC_MF_mVc_VAR) %in% "Type")])

# Special mMSE elements
SP_All_mMSE_Bias<-rbind(OSMAC_mMSE_Bias[,!(colnames(OSMAC_mMSE_Bias) %in% "Type")],
                        OSMAC_MF_mMSE_Bias[,!(colnames(OSMAC_MF_mMSE_Bias) %in% "Type")])
SP_All_mMSE_EMSE<-rbind(OSMAC_mMSE_EMSE[,!(colnames(OSMAC_mMSE_EMSE) %in% "Type")],
                        OSMAC_MF_mMSE_EMSE[,!(colnames(OSMAC_MF_mMSE_EMSE) %in% "Type")])
SP_All_mMSE_MSE<-rbind(OSMAC_mMSE_MSE[,!(colnames(OSMAC_mMSE_MSE) %in% "Type")],
                       OSMAC_MF_mMSE_MSE[,!(colnames(OSMAC_MF_mMSE_MSE) %in% "Type")])
SP_All_mMSE_Info<-rbind(OSMAC_mMSE_Info[,!(colnames(OSMAC_mMSE_Info) %in% "Type")],
                        OSMAC_MF_mMSE_Info[,!(colnames(OSMAC_MF_mMSE_Info) %in% "Type")])
SP_All_mMSE_VAR<-rbind(OSMAC_mMSE_VAR[,!(colnames(OSMAC_mMSE_VAR) %in% "Type")],
                       OSMAC_MF_mMSE_VAR[,!(colnames(OSMAC_MF_mMSE_VAR) %in% "Type")])

# Special mVc elements
SP_All_mVc_Bias<-rbind(OSMAC_mVc_Bias[,!(colnames(OSMAC_mVc_Bias) %in% "Type")],
                        OSMAC_MF_mVc_Bias[,!(colnames(OSMAC_MF_mVc_Bias) %in% "Type")])
SP_All_mVc_EMSE<-rbind(OSMAC_mVc_EMSE[,!(colnames(OSMAC_mVc_EMSE) %in% "Type")],
                        OSMAC_MF_mVc_EMSE[,!(colnames(OSMAC_MF_mVc_EMSE) %in% "Type")])
SP_All_mVc_MSE<-rbind(OSMAC_mVc_MSE[,!(colnames(OSMAC_mVc_MSE) %in% "Type")],
                       OSMAC_MF_mVc_MSE[,!(colnames(OSMAC_MF_mVc_MSE) %in% "Type")])
SP_All_mVc_Info<-rbind(OSMAC_mVc_Info[,!(colnames(OSMAC_mVc_Info) %in% "Type")],
                        OSMAC_MF_mVc_Info[,!(colnames(OSMAC_MF_mVc_Info) %in% "Type")])
SP_All_mVc_VAR<-rbind(OSMAC_mVc_VAR[,!(colnames(OSMAC_mVc_VAR) %in% "Type")],
                       OSMAC_MF_mVc_VAR[,!(colnames(OSMAC_MF_mVc_VAR) %in% "Type")])

All_Objects<-ls()
remove(list = All_Objects[startsWith(All_Objects,"OSMAC")])
remove(list = All_Objects[startsWith(All_Objects,"RS")])
```

# Best Sub-sampling Method : `r params$Model_Path`

# Compare SMSE, EMSE, Variance, Information and Bias for all methods.

```{r Compare All,fig.width=20,fig.height=12}
All_Bias %>% 
  ggplot(.,aes(x=Subsample_Size,y=mean_Bias,color=Model))+
  geom_point()+geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("Bias")+ggtitle("Bias")->p1

All_MSE %>%
  ggplot(.,aes(x=Subsample_Size,y=MSE,color=Model))+
  geom_point()+geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("MSE")+ggtitle("MSE")->p2

All_EMSE %>%
  ggplot(.,aes(x=Subsample_Size,y=EMSE,color=Model))+
  geom_point()+geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("EMSE")+ggtitle("EMSE")->p3

All_VAR %>%
  ggplot(.,aes(x=Subsample_Size,y=mean_Variance,color=Model))+
  geom_point()+geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("Variance")+ggtitle("Variance")->p4

All_Info %>%
  ggplot(.,aes(x=Subsample_Size,y=mean_Info,color=Model))+
  geom_point()+geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("Information")+ggtitle("Information")->p5

ggarrange(p1,p2,p3,p4,p5,common.legend = TRUE,nrow = 2,ncol = 3)
```

# Special Comparison of SMSE, EMSE, Variance, Information and Bias for mMSE.
## (Real Model and Assumed Model results with Full Data Real Model)

```{r Special Compare 1,fig.width=20,fig.height=12}
SP_All_mMSE_Bias %>%
  ggplot(.,aes(x=Subsample_Size,y=mean_Bias,color=Model))+
  geom_point()+geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("Bias")+ggtitle("Bias")->p1

SP_All_mMSE_MSE %>%
  ggplot(.,aes(x=Subsample_Size,y=MSE,color=Model))+
  geom_point()+geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("MSE")+ggtitle("MSE")->p2

SP_All_mMSE_EMSE %>%
  ggplot(.,aes(x=Subsample_Size,y=EMSE,color=Model))+
  geom_point()+geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("EMSE")+ggtitle("EMSE")->p3

SP_All_mMSE_VAR %>%
  ggplot(.,aes(x=Subsample_Size,y=mean_Variance,color=Model))+
  geom_point()+geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("Variance")+ggtitle("Variance")->p4

SP_All_mMSE_Info %>%
  ggplot(.,aes(x=Subsample_Size,y=mean_Info,color=Model))+
  geom_point()+geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("Information")+ggtitle("Information")->p5

ggarrange(p1,p2,p3,p4,p5,common.legend = TRUE,nrow = 2,ncol = 3)
```

# Compare  SMSE, EMSE, Variance, Information and Bias for mVc.
## (Real Model and Assumed Model results with Full Data Real Model)

```{r Special Compare 2,fig.width=20,fig.height=12}
SP_All_mVc_Bias %>%
  ggplot(.,aes(x=Subsample_Size,y=mean_Bias,color=Model))+
  geom_point()+geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("Bias")+ggtitle("Bias")->p1

SP_All_mVc_MSE %>%
  ggplot(.,aes(x=Subsample_Size,y=MSE,color=Model))+
  geom_point()+geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("MSE")+ggtitle("MSE")->p2

SP_All_mVc_EMSE %>%
  ggplot(.,aes(x=Subsample_Size,y=EMSE,color=Model))+
  geom_point()+geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("EMSE")+ggtitle("EMSE")->p3

SP_All_mVc_VAR %>%
  ggplot(.,aes(x=Subsample_Size,y=mean_Variance,color=Model))+
  geom_point()+geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("Variance")+ggtitle("Variance")->p4

SP_All_mVc_Info %>%
  ggplot(.,aes(x=Subsample_Size,y=mean_Info,color=Model))+
  geom_point()+geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+
  ylab("Information")+ggtitle("Information")->p5

ggarrange(p1,p2,p3,p4,p5,common.legend = TRUE,nrow = 2,ncol = 3)
```